---
title: "ETC5521 Assignment 1"
subtitle: "Australian Animal Outcomes"
team: galah
author:
  - Sen Wang
  - Muhammad Soban Qasim
date: "`r Sys.Date()`"
bibliography: references.bib
biblio-style: authoryear-comp
output: 
  bookdown::html_document2
---

[This assignment is for ETC5521 Assignment 1 by Team `r rmarkdown::metadata$team` comprising of `r knitr::combine_words(rmarkdown::metadata$author)`.]{style="color:#006DAE;"}


# Introduction and motivation

Animals are all around us in the sky, in the oceans, in our backyards and in our homes. What happens to animals in trouble, it could be cats, dogs, horses or livestock? Who takes care of them when they are in trouble or are lost? The answer is Royal Society for the Prevention of cruelty to Animals.  

The RSPCA is Australia's oldest, largest and most trusted animal welfare organization, according to its description, there is a gradual increase in the number of animals every year which are received in animal shelters and adoption centres across the country. RSPCA decides to analyse the results of animals in Australia and help them.

While looking online and researching on RSPCA we came across many articles that were indirectly blaming RSPCA for their high euthanize rates and some individuals who saw the worst side of RSPCA. That got us interested in the outcomes from RSPCA and as RSPCA operates state wise. After a while it raised a question which animal is euthanized most.


[https://mobile.abc.net.au/news/2020-02-26/rscpa-qld-issues-apology-after-accidentally-euthanasing-kitten/12001668]: here

We also came across multiple articles and web pages dedicated to adoption of animals that have been in the care of RSPCA. We wanted to know which state rehomes the most animals. 

Finally, cat person or dog person a comparison between the 2 types of people out there, which lost pet is most likely to get reclaimed.

## Limitations



# Data description

There are three data tables separately coming from the RSPCA, Townsville Animal Complaints and Brisbane Open Data - Animal Complaints.

1. animal_outcomes.csv

*year*: Full year
*animal_type*: Animal type (horse, wildlife, dog, cat, etc)
*outcome*: 	Animal outcome - euthanized, released, rehomed, etc
*ACT*: ACT - Australian Capital Territory
*NSW*: New South Wales
*NT*: Northern Territory
*QLD*: Queensland
*SA*: South Australia
*TAS*: Tasmania
*VIC*: Victoria
*WA*: Western Australian
*Total*: Australian Total

2. animal_complaints.csv

*Animal Type*: Animal Type
*Complaint Type*: Complaint type
*Date Received*: Date received (Month - year)
*Suburb*: Suburb/region
*Electoral Division*: Electoral Division

3. brisbane_complaints.csv

*nature*:	Nature of complaints (animal)
*animal_type*: Animal type
*category*: Category of complaint
*suburb*: Suburb where reported
*date_range*: Date range (typically 1 quarter + year)
*responsible_office*: Responsible office for the complaint
*city*: City (Brisbane)




 

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = F, message = FALSE, warning = FALSE)
```

```{r libraries}
library(tidytuesdayR)
library(tidyr)
library(ggplot2)
library(dplyr)
library(rgdal)
library(maptools)
library(plyr)
library(rgeos)
library(plotly)
library(mapproj)
library(leaflet)
library(tmaptools)
library(sf)
library(broom)
library(kableExtra)
```

```{r load-data, echo=FALSE, message=FALSE}

tuesdata <- tidytuesdayR::tt_load(2020, week = 30) #downloading the data set as instructed in the git repo

animal_outcomes <- tuesdata$animal_outcomes  #seperating the data files for easier navigation
animal_complaints <- tuesdata$animal_complaints
brisbane_complaints <- tuesdata$brisbane_complaints

```




# Questions


## Q1: Which state has highest rate for the best possible outcome for animals?

```{r loading rds map}

aus_map <- readRDS("oz_map.rds") #loading the rds map file 

```

```{r calculating good outcomes}

q1 <- animal_outcomes %>% select(-year, 
                   -animal_type) %>% 
  group_by(outcome) %>% 
  summarise_at(vars(ACT, 
                    NSW, 
                    NT, 
                    QLD, 
                    SA, 
                    TAS, 
                    VIC, 
                    WA),  
               list( sum),na.rm=TRUE) #suming up the outcomes category wise




q11 <-  pivot_longer(q1, `ACT`:`WA`, 
                     names_to = "State", 
                     values_to = "outcomes") %>% 
  pivot_wider(names_from = outcome, 
              values_from = outcomes) #wranggling to get better visualizing sense

q11 <- q11 %>%  mutate(good_outcome = (rowSums(q11[,c(-1,-3)]))/rowSums(q11[,c(-1)])*100) #calculating percentage of good outcomes by taking out euthanization and dividing it by the total number of outcomes per state 

                                                                                                    q12 <- q11 %>% select(State, 
                                                                                                                         good_outcome)           #selecting variables for ease 
                                                                                                    
                                                                                                    
```


```{r getting the states up}
aus_map_states <- aus_map %>% 
  left_join(q12, by = c("state_abb" = "State")) #joining data sets by states

```


```{r Animalsurvivalrate, fig.height=8, fig.width=10,fig.cap="Australia Euthanization rate", message=FALSE, warning=FALSE}
state_plot <- ggplot(aus_map_states, 
                     aes(x = longitude, 
                         y = latitude, 
                         fill = good_outcome)) + 
  geom_sf() + aes(x = longitude, 
                  y = latitude) +   
  labs(subtitle="Animals saved by RSPCA", 
       y=" ", 
       x=" ", 
       title="Animal survival rate", 
       caption = "Australia Euthanization rate")+ 
  geom_sf_text(aes(label = state), 
               colour = "white")
  scale_fill_viridis_c( "alpha = 1",
  "begin = 0",
  "end = 100",
  "direction = 1",
  option = "D",
  values = NULL,
  space = "Lab",
  na.value = "black",
  guide = "colourbar",
  aesthetics = "fill"
) 
  
state_plot
```




Figure  \@ref(fig:Animalsurvivalrate), shows the rate of animals that are in the care of RSPCA, find a good outcome, from the year 1999 to 2018, where animals can be either reclaimed, rehomed, released, in stock, transferred or be currently in care. The plot clearly shows Western Australia to have the highest good outcome rate of over 75%, while sadly New South Wales has the lowest good outcome rate of 55%. Otherwise meaning that the rest of the 45% of animals are euthanized.  






## Q2: Which animal is euthanized the most throughout Australia?



```{r getting counts, include=FALSE}
animal_outcomes  %>% group_by( year,
                 outcome, 
                 animal_type) %>%                         
  summarise_at(vars(ACT, 
                    NSW, 
                    NT, 
                    QLD, 
                    SA, 
                    TAS, 
                    VIC),  
               list( sum)) 

q2 <- animal_outcomes %>% filter(outcome == "Euthanized") %>% 
  select(animal_type, 
         ACT, 
         NSW, 
         NT, 
         QLD, 
         SA, 
         TAS, 
         VIC, 
         WA) %>%  
  group_by(animal_type) %>%                         
  summarise_at(vars(ACT, 
                    NSW, 
                    NT, 
                    QLD, 
                    SA, 
                    TAS, 
                    VIC, 
                    WA),            
               list(" " = sum)) 

 q2$total_animals_euthanised = rowSums(q2[,c(-1)], na.rm=TRUE)
```

```{r Australia Euthanization count, fig.width=10}
options(scipen=999)
 p2 <- ggplot(q2,
              aes(x =reorder(animal_type, 
                             -total_animals_euthanised), 
                  y =  total_animals_euthanised)) + 
   geom_bar(stat="identity", 
            fill="tomato3") +   
   labs(subtitle="Which animal is the most euthanized", 
       y="Count", 
       x="Animals", 
       title="Australia Euthanization count", 
       caption = "Australia Euthanization rate") + 
  theme(axis.text.x = element_text(angle=45, vjust=0.9)) + 
   ylim(c(0, 600000)) 

ggplotly(p2)
```

Figure  \@ref(fig:Australia Euthanization count), shows us the euthanization count by animal type throughout Australia from the year 1999 to 2018. Cats have the highest recorded euthanization count at 553,956 in Australia under the care of RSPCA. 


## Q3: What is the percentage of rehomed animals state wise?

```{r Rehomed}
animal_outcomes2 <- animal_outcomes %>%
  filter(outcome == "Rehomed")


ACT_Rehomed_Ttl <- sum(animal_outcomes2$ACT)
NSW_Rehomed_Ttl <- sum(animal_outcomes2$NSW)
NT_Rehomed_Ttl <- sum(animal_outcomes2$NT, na.rm = TRUE)
QLD_Rehomed_Ttl <- sum(animal_outcomes2$QLD)
SA_Rehomed_Ttl <- sum(animal_outcomes2$SA)
TAS_Rehomed_Ttl <- sum(animal_outcomes2$TAS)
VIC_Rehomed_Ttl <- sum(animal_outcomes2$VIC)
WA_Rehomed_Ttl <- sum(animal_outcomes2$WA)
```

```{r Total}
ACT_Ttl <- sum(animal_outcomes$ACT)
NSW_Ttl <- sum(animal_outcomes$NSW)
NT_Ttl <- sum(animal_outcomes$NT, na.rm = TRUE)
QLD_Ttl <- sum(animal_outcomes$QLD)
SA_Ttl <- sum(animal_outcomes$SA, na.rm = TRUE)
TAS_Ttl <- sum(animal_outcomes$TAS, na.rm = TRUE)
VIC_Ttl <- sum(animal_outcomes$VIC)
WA_Ttl <- sum(animal_outcomes$WA)
```

```{r New_dataset}
d <- data.frame(
  State = c("ACT", 
            "NSW", 
            "NT", 
            "QLD", 
            "SA", 
            "TAS", 
            "VIC", 
            "WA"),
  Rehomed_Ttl = c(ACT_Rehomed_Ttl, 
                  NSW_Rehomed_Ttl, 
                  NT_Rehomed_Ttl, 
                  QLD_Rehomed_Ttl, 
                  SA_Rehomed_Ttl, 
                  TAS_Rehomed_Ttl, 
                  VIC_Rehomed_Ttl, 
                  WA_Rehomed_Ttl),
  Total = c(ACT_Ttl, 
            NSW_Ttl, 
            NT_Ttl, 
            QLD_Ttl, 
            SA_Ttl, 
            TAS_Ttl, 
            VIC_Ttl, 
            WA_Ttl)
)  %>%
  mutate(Percent = round(Rehomed_Ttl/Total*100))
```


```{r p1, fig.width = 8, fig.height = 8, fig.cap="The percentage of rehomed animals state wise", fig.align= "center"}
p1 <- ggplot(d, 
             aes(x=reorder(State,Percent), 
                 y=Percent)) +
  geom_histogram(stat="identity", 
                 fill = "lightblue", 
                 position = "dodge",
                 width= 0.8) +
  xlab("State") +
  ylab("Percentage") +
  ggtitle("The percentage of rehomed animals state wise")

ggplotly(p1)
```

In Figure  \@ref(fig:p1), it is clear that the highest proportion of animals were relocated in WA and the lowest proportion in QLD, but the difference between them is so much as only 13 percent. This also corroborates with Figure  \@ref(fig:Animal survival rate) in terms of good outcome. 





## Q4: Cats Vs Dogs who are more likely to get reclaimed?
 




```{r screening}
animal_outcomes3 <- animal_outcomes %>%
  filter(outcome == "Reclaimed") %>%
  filter(animal_type == c("Dogs", "Cats"))
```

```{r p2, fig.width = 8, fig.height = 8, fig.cap="The more likely to get reclaimed", fig.align= "center"}
p2 <-  ggplot(animal_outcomes3, 
              aes( x = year,
              y = Total, 
              group = as.factor(animal_type),
              shape = as.factor(animal_type),
              colour = as.factor(animal_type))) +
  geom_line() +
  geom_point() +
  ggtitle("The more likely to get reclaimed") +
  theme(legend.title=element_blank())

ggplotly(p2)
```

In Figure  \@ref(fig:p2), from 1999 to 2018 the number of reclaimed cats has been stable, while the number of reclaimed dogs has fluctuated over the years. 2009 was a good year for dogs.

### While we are at the subject of Cats Vs Dogs 

```{r complaints}
x1 <- animal_complaints %>% 
  select(`Animal Type`) %>% 
  count() 

```

```{r Complaints against who Cats or Dogs}
p4e <- x1 %>% ggplot(aes(x = Animal.Type, 
                 y = freq)) + 
  geom_bar(stat="identity", 
           fill="yellow") + 
  ylab("complaint_count") + xlab("cat/Vs/dog")

ggplotly(p4e)
```

Figure  \@ref(fig:Complaints against who Cats or Dogs), shows that dog complaints made by people are disproportionately higher than people who have filed a complaint against cats.


# Analysis and findings

[FILL] Should include at least one plot or numerical summary for each of your questions, that helps the reader arrive at an answer. You should also write paragraphs describing the methods, summaries and findings. 




# References
